Spin canonical divisors of log stacky curves

### 2015-06-21 stacky_write_up/spin_cring.tex
added missing L's in base cases
adding corollaries to section 3

Then we find $2\halfcan=D+\sum \subhalf{i}P_i$ where $D\sim K_X+
\Delta\sim -2\infty+ \Delta$ is a divisor of $X$ (i.e. with 
possibly negative integer coefficients).  Since $\Delta$ is non-
zero effective, $\deg(D)\ge -1$.   Noting that the coefficient of 
any point $P$ occurring in a stacky divisor of $\sx$ must have 
coefficient lying in $\mathbf{Z}[\frac{1}{e_i}]$ (ranging over $e_i$
s appearing in the characteristic of $(\sx,\Delta)$).

Since all $e_i$s are odd, $2$ cannot appear in a denominator of the 
coefficient of a point $P$ in $\halfcan$, meaning that $\frac{D}{2}$
 must be a $X$-divisor.  Since $\deg(\frac{D}{2})=\frac{1}{2}\deg(D)
\ge -\frac{1}{2}$, we in fact have $\deg(D)\ge 0$.  Thus $\halfcan$ 
is linearly equivalent to an effective $X$ divisor plus $\sum 
\subhalf{i}P_i$.  We have now reduced to the case of $\halfcan$ 
linearly equivalent to an effective divisor.  The result in this 
specific case in fact generalizes to any $\mathbf{Q}$-divisor: $D=\sum
\frac{\alpha_i}{\beta_i} P_i$ for $\alpha_i,\beta_i\in \mathbf{N}$.

@@

\begin{align*}
	\initial_\prec(I') &= \initial_\prec(I) k[u, x_1, \ldots, x_m, y_1, \ldots, y_n] \\
										 &+ \langle y_i x_j: 1 \le i \le n, 1 \le j \le m \rangle \\
										 &+ \langle y_i y_j: \lfloor \frac{\alpha}{\beta} \rfloor \rangle \le i \le j \le n-1 \\
										 &+ \langle y_i y_j: 1\le i\le j\lfloor \frac{\alpha}{\beta} \rfloor\\
										 &+ \langle (y_i)^{\deg(y_{i+1})}: 1\le i=\lfloor \frac{\alpha}{\beta} \rfloor
\end{align*}
In particular, if $\frac{\alpha}{\beta}\le 1$, then $I$ is minimally generated by elements with leading terms of the form $y_ix_j$ and $y_iy_j$.
\item If $\gin_\prec(I)=\initial_\prec(I)$ then $\gin_\prec(I')=\initial_\prec(I')$.
\end{itemize}
\end{lem}

@@

Let $J'$ be the ideal of $k[u, x_1, \ldots, x_m, y_1, \ldots, y_n]$ generated by the initial 
terms described in the previous paragraphs together with $\gin_\prec(I)k[u, x_1, \ldots, 
x_m, y_1, \ldots, y_n]$.  We show that $J$ is in fact all of the initial ideal.  Suppose $f$ is a 
non-constant monomial that does not lie within $J$.  Further suppose $f$ does not lie in 
$k[u, x_1, \ldots, x_m]$.  Then there is some $y_i|f$.  Since $f\not\in j$, we cannot have 
some $x_j|f$ or else we would have $y_ix_j|f$ contradicting $f\not\in J$.  Furthermore, if 
$y_l|f$ for $l<i$, then we must have $d_l=1$ or else $y_ly_i$ is an initial element of a 
relation and divides $f$.  If $d_i=1$ as well, then we must have $c_i=\lfloor \frac{\alpha}
{\beta}\rfloor$, or else $y_iy_l\in I$; finally we can not have $(y_{\lfloor \frac{\alpha}{\beta}
\rfloor})^{d_{\lfloor \frac{\alpha}{\beta} \rfloor+1}+1}|f$.  We observe that the remaining 
elements are precisely those of the form $y_n^a y_i u^b$, $y_n^a zy_iu^b$, and $y_n^a z 
u^b$ which are the elements of our chosen $k$-basis for $k[u, x_1, \ldots, x_m, y_1, \ldots, 
y_n]/I'$ over $k[u, x_1, \ldots, x_m]/I$.  Therefore, by dimension counting (in each degree), 
we see that $J'$ is the initial ideal of $I'$.  Furthermore, since our choice of $y_i$s was 
generic, this is in fact the generic initial ideal.
\end{proof}

@@

\begin{cor}

\end{cor}

@@

	\hline

@@

\begin{enumerate}
	\item[(Ad-i)] Suppose for all $i \in J$, there exists $y_{i,e_i'} 
		\in R'$ so that $R'$ admits a presentation
		\begin{align*}
			R' \cong \left( k[x] \otimes k[y_{i,e_i'}]_{i \in J} \right)/I'
		\end{align*}
		so that for all $i \in J$ we have $\deg y_{e_i'} = e_i'$ and $-ord_{
		Q_i}(y_{i,e_i'}) = \frac{e_i'-1}{2}$.
	\item[(Ad-ii)] For all $i \in J$, any generator $z \neq y_{i,e_i'}$
		of $R'$ satisfies
		\begin{align*}
			\frac{- ord_{Q_i}(z)}{\deg z} < \subhalf {e_i'}.
		\end{align*}
	\item[(Ad-iii)] For all $i \in J$ we have
		\begin{align*}
			\deg \lfloor e_i L \rfloor \geq 2g - 1 + \max_{d \geq 0} \# S_{\sigma, J}(i, d)
		\end{align*}
		\todo{Check the 2g - 1, in particular, David seems to think we are usually 1 off and possibly 2 off when g = 1. I have a feeling this is a mistake, but don't understand why.} 
		where
		\begin{align*}
			S_{\sigma, J}(i, d) := \{j \in J : j \neq i \text{ and } e_j'+2d
			\mid e_i + 2d\}
		\end{align*}
\end{enumerate}
\end{defn}

@@

\begin{enumerate}
	\item[(a)] $\sigma = (0; 3, 3, 11; 0), J = \{3\}$
	\item[(b)] $\sigma = (0; 3, 5, 9; 0), J = \{2, 3\}$
	\item[(c)] $\sigma = (0; 5, 5, 7; 0), J = \{1, 2, 3\}$
	\item[(d)] $\sigma = (0; 3, 3, 3, 5; 0), J = \{1, 2, 3, 4\}$
	\item[(e)] $\sigma = (0; 3, 3, 3, 3, 3; 0), J = \{1, 2, 3, 4, 5\}$
\end{enumerate}
\end{lem}

@@

\begin{enumerate}
	\item[(a)] $(\deg \lfloor e_3 \rfloor, \# S_{(\sigma, J)}(3)) = (1, 0)$

	\item[(b)] $(\deg \lfloor e_2 \rfloor, \#S_{(\sigma, J)}(2)) = (1, 0)$ \\
		$(\deg \lfloor e_3 \rfloor, \#S_{(\sigma, J)}(3)) = (1, 0)$

	\item[(c)] $(\deg \lfloor e_1 \rfloor, \#S_{(\sigma, J)}(1)) = (2, 1)$ \\
		$(\deg \lfloor e_2 \rfloor, \#S_{(\sigma, J)}(2)) = (2, 1)$ \\
		$(\deg \lfloor e_3 \rfloor, \#S_{(\sigma, J)}(3)) = (1, 0)$

	\item[(d)] $(\deg \lfloor e_i \rfloor, \#S_{(\sigma, J)}(i)) = (3, 1)$
		for $i \neq 4$ \\
		$(\deg \lfloor e_4 \rfloor, \#S_{(\sigma, J)}(4))	= (2, 0)$.

	\item[(e)] $(\deg \lfloor e_i \rfloor, \#S_{(\sigma, J)}(1)) = (5, 0)$
		for all $i \in J$.
\end{enumerate}

>>>>

@@

\begin{align*}
	\initial_\prec(I') &= \initial_\prec(I) k[u, x_1, \ldots, x_m, y_1, \ldots, y_n] \\
										 &+ \langle y_i x_j: 1 \le i \le n, 1 \le j \le m \rangle \\
										 &+ \langle y_i y_j: \lfloor \frac{\alpha}{\beta} \rfloor \rangle \le i \le j \le n-1 \\
										 &+ \langle y_i y_j: 1\le i\le j\lfloor \frac{\alpha}{\beta} \rfloor\\
										 &+ \langle (y_i)^{\deg(y_{i+1}+1)}: 1\le i=\lfloor \frac{\alpha}{\beta} \rfloor
\end{align*}
In particular, if $\frac{\alpha}{\beta}\le 1$, then $I$ is minimally generated by elements with leading terms of the form $y_ix_j$ and $y_iy_j$.
\item If $\gin_\prec(I)=\initial_\prec(I)$ then $\gin_\prec(I')=\initial_\prec(I')$.
\item A Grobner basis with the initial terms $y_i x_j$, $y_iy_j$, and possibly $(y_{\lfloor \frac{\alpha}{\beta} \rfloor})^{y_{\lfloor \frac{\alpha}{\beta} \rfloor}+1}$ as shown above minimally generates $I'$.
\end{itemize}
\end{lem}

@@

Let $J'$ be the ideal of $k[u, x_1, \ldots, x_m, y_1, \ldots, y_n]$ generated by the initial 
terms described in the previous paragraphs together with $\gin_\prec(I)k[u, x_1, \ldots, 
x_m, y_1, \ldots, y_n]$.  We show that $J$ is in fact all of the initial ideal.  Suppose $f$ is a 
non-constant monomial that does not lie within $J$.  Further suppose $f$ does not lie in 
$k[u, x_1, \ldots, x_m]$.  Then there is some $y_i|f$.  Since $f\not\in j$, we cannot have 
some $x_j|f$ or else we would have $y_ix_j|f$ contradicting $f\not\in J$.  Furthermore, if 
$y_l|f$ for $l<i$, then we must have $d_l=1$ or else $y_ly_i$ is an initial element of a 
relation and divides $f$.  If $d_i=1$ as well, then we must have $c_i=\lfloor \frac{\alpha}
{\beta}\rfloor$, or else $y_iy_l\in I$; finally we can not have $(y_{\lfloor \frac{\alpha}{\beta}
\rfloor})^{d_{\lfloor \frac{\alpha}{\beta} \rfloor+1}+1}|f$.  We observe that the remaining 
elements are precisely those of the form $y_n^a y_i u^b$, $y_n^a zy_iu^b$, and $y_n^a z 
u^b$ which are the elements of our chosen $k$-basis for $k[u, x_1, \ldots, x_m, y_1, \ldots, 
y_n]/I'$ over $k[u, x_1, \ldots, x_m]/I$.  Therefore, by dimension counting (in each degree), 
we see that $J'$ is the initial ideal of $I'$.  Furthermore, since our choice of $y_i$s was 
generic, this is in fact the generic initial ideal.

\todo{Extend Lemma \ref{lem:minimal:quadratic} to commutative diagrams to show minimality}
\end{proof}

\begin{rem}\label{rem:deg1_sat_ind_gen_rel_degrees}
Notice that in lemma \ref{lem:deg1_sat_ind}, $D'$ is generated in degree at most $\max(\delta,\beta)$ with relations generated in degree at most $\max(\tau,\delta+\beta,2\beta)$
\end{rem}

@@

\begin{cor}\label{cor:effective_Q_divisor_can_ring}
If the genus of $X$ is 0 and $D\sim\sum \frac{\alpha_i}{\beta_i} P_i$ is linearly equivalent to an effective $\mathbf{Q}$ divisor of $X$, then $R_D$ is generated in degree at most $\max(\beta_i,1)$ with relations generated in degree at most $2\max(\beta_i,1)$.
\end{cor}

\begin{proof}
We can prove this by induction.  As a base case, let $D\sim 0$.  Then $R_D\cong k[x]$ which is generated in degree 1 with no relations, satisfying the inductive hypothesis.
Now, suppose the result is proven for all effective $D$ with support at most $r$ points; then if $D'\sim \sum_{i=1}^{r+1} \frac{\alpha_i}{\beta_i}$, we set $D=\sum_i^{r} \frac{\alpha_i}{\beta_i}$.  By hypothesis, $R_D$ is generated in degree at most $\max_{1\le i\le r}(\beta_i)$ with relations generated in degree at most 2$\max_{1\le i\le r}(\beta_i)$.  Since $D$ is effective, the condition of equation \ref{eqn:deg1_sat_ind_dim} is met, so by remark \ref{rem:deg1_sat_ind_gen_rel_degrees} of lemma \ref{lem:deg1_sat_ind}, $R_{D'}$ is generated in degree at most $\max(\beta_i)$ with relations generated in degree at most $2\max(\beta_i)$.
\end{proof}

\begin{cor}\label{cor:genus_0_posiive_delta}
Let $\halfcan$ be a stacky log spin canonical divisor of signature $(0,e_1,...,e_n,\delta)$ with $\delta>0$.  Then 
\end{cor}
\begin{proof}
Since $\halfcan$ is a half canonical divisor, 
\[
	2\halfcan=D+\sum \subhalf{i}P_i
\] 
where 
\[
	D\sim K_X+\Delta\sim -2\infty+ \Delta
\] is a divisor of $X$ (i.e. with 
possibly negative integer coefficients).  Since $\Delta$ is non-
zero effective, $\deg(D)\ge -1$.   Noting that the coefficient of 
any point $P$ occurring in a stacky divisor of $\sx$ must have 
coefficient lying in $\mathbf{Z}[\frac{1}{e_i}]$ (ranging over $e_i$
s appearing in the characteristic of $(\sx,\Delta)$).

Since all $e_i$s are odd, $2$ cannot appear in a denominator of the 
coefficient of a point $P$ in $\halfcan$, meaning that $\frac{D}{2}$
 must be a $X$-divisor.  Since $\deg(\frac{D}{2})=\frac{1}{2}\deg(D)
\ge -\frac{1}{2}$, we in fact have $\deg(D)\ge 0$.  Thus $\halfcan$ 
is linearly equivalent to an effective $X$ divisor plus $\sum 
\subhalf{i}P_i$.  We have thus shown $\halfcan$ is linearly equivalent to an effective divisor.  Thus, by \ref{cor:effective_Q_divisor_can_ring} the desired result follows.
\end{proof}

\begin{cor}
Let $D$ be a fractional $\frac{a}{b}$-canonical divisor of a genus 0 stacky curve with denominator $b > 2$ when $\frac{a}{b}$ is written in reduced form.
\end{cor}
\begin{proof}
For a similar reason as in corollary \ref{cor:genus_0_posiive_delta}, $D$ must be equivalent to an effective divisor (in this case even when $\delta=0$), so by corollary \ref{cor:effective_Q_divisor_can_ring} we get the desired result.
\end{proof}

\todo{Add higher genus corollaries once we figure out what we want to say there}

@@

	\hline
	\hline

@@

\begin{enumerate}
	\item[(Ad-i)] Suppose for all $i \in J$, there exists $y_{i,e_i'} 
		\in R'$ so that $R'$ admits a presentation
		\begin{align*}
			R' \cong \left( k[x] \otimes k[y_{i,e_i'}]_{i \in J} \right)/I'
		\end{align*}
		so that for all $i \in J$ we have $\deg y_{e_i'} = e_i'$ and $-ord_{
		Q_i}(y_{i,e_i'}) = \frac{e_i'-1}{2}$.
	\item[(Ad-ii)] For all $i \in J$, any generator $z \neq y_{i,e_i'}$
		of $R'$ satisfies
		\begin{align*}
			\frac{- ord_{Q_i}(z)}{\deg z} < \subhalf {e_i'}.
		\end{align*}
	\item[(Ad-iii)] For all $i \in J$ we have
		\begin{align*}
			\deg \lfloor e_i L \rfloor \geq 2g - 1 + \max_{d \geq 0} \# S_{\sigma, J}(i, d)
		\end{align*}
		\todo{Check the 2g - 1, in particular, David seems to think we are usually 1 off and possibly 2 off when g = 1. I have a feeling this is a mistake, but don't understand why.} 
		where
		\begin{align*}
			S_{\sigma, J}(i, d) := \{j \in J : j \neq i \text{ and } e_j'+2d
			\mid e_i - e_j'\}
		\end{align*}
\end{enumerate}
\end{defn}

@@

\begin{enumerate}
	\item[(a)] $\sigma_1 = (0; 3, 3, 11; 0), J_1 = \{3\}$
	\item[(b)] $\sigma_2 = (0; 3, 5, 9; 0), J_2 = \{2, 3\}$
	\item[(c)] $\sigma_3 = (0; 5, 5, 7; 0), J_3 = \{1, 2, 3\}$
	\item[(d)] $\sigma_4 = (0; 3, 3, 3, 5; 0), J_4 = \{1, 2, 3, 4\}$
	\item[(e)] $\sigma_5 = (0; 3, 3, 3, 3, 3; 0), J_5 = \{1, 2, 3, 4, 5\}$
\end{enumerate}
\end{lem}

@@

\begin{enumerate}
	\item[(a)] $(\deg \lfloor e_3 \rfloor, \max_{d \geq 0} \# S_{(
		\sigma_1, J_1)}(3)) = (1, 0)$

	\item[(b)] $(\deg \lfloor e_2 \rfloor, \max_{d \geq 0} \#S_{(
		\sigma_2, J_2)}(2)) = (1, 0)$ \\
		$(\deg \lfloor e_3 \rfloor, \max_{d \geq 0} \#S_{(\sigma_2, J_2)}(3))
		= (1, 0)$

	\item[(c)] $(\deg \lfloor e_1 \rfloor, \max_{d \geq 0}
		\#S_{(\sigma_3, J_3)}(1)) = (2, 1)$ \\
		$(\deg \lfloor e_2 \rfloor, \max_{d \geq 0}
		\#S_{(\sigma_3, J_3)}(2)) = (2, 1)$ \\
		$(\deg \lfloor e_3 \rfloor, \max_{d \geq 0}
		\#S_{(\sigma_3, J_3)}(3)) = (1, 0)$

	\item[(d)] $(\deg \lfloor e_i \rfloor, \max_{d \geq 0}
		\#S_{(\sigma_4, J_4)}(i)) = (3, 1)$	for $i \neq 4$ \\
		$(\deg \lfloor e_4 \rfloor, \max_{d \geq 0} \#S_{(\sigma_4, J_4)}(4))
		= (2, 0)$.

	\item[(e)] $(\deg \lfloor e_i \rfloor, \max_{d \geq 0}
		\#S_{(\sigma_5, J_5)}(1)) = (5, 0)$	for all $i \in J$.
\end{enumerate}
